---
title: "Portfolio 1: Math-ing in R"
---

> This may come as a shock to no one, but R can do math. In this portfolio piece, I will demonstrate how to do some math in R that I find particularly enjoyable. 

Let's load the relevant packages first, among these are the package mosaiCalc (CITATION NEEDED) which will do some of the heavy lifting.

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
library(mosaicCalc)
#this is also helpful and necessary 
help(package="mosaicCalc")
```

We will start with vectors, something all of us can get behind. Contrasts are vectors, and we love contrasts, don't we? 
```{r}
#first making a simple linear combination 
#add in an actual example from Eric's class, show that they dot to zero using R 
project()

```

To plot a vector field in R, we can use the vectorfield_plot function. It seems like it can only handle two dimensional vectors but I will certainly circle back with a more interesting example if I learn otherwise. The use of tilde expressions are terrible, but my example functions are ..... 

```{r}
#plot a vector field 
vectorfield_plot(x ~ tan(y), y ~ tan(x), bounds(x=-1:1, y=-1:1),
  npts = 5,
  color = "pink",
  alpha = 0.9,
  transform = sqrt,
  env = NULL
)
  
```
As it turns out, this particular method will not handle three variables, but apparently one strategy is to use heatmaps to visualize the third variable. That is something to play around with in the future. 

I'm bored of this, so I am going to practice taking some plain old boring integrals. First thing to note is that ~ is the enemy. They show up under lots of different conditions. At least in the context of differentiation, our friend "~" means "with respect to," which is fine. 

```{r}
#Tutorial Example, need to come up with my own stuff - I will just run through pauls online notes problems in here and time if I can do it faster by hand vs on R 
library(pracma) #for the integrate function, because the functions in mosiacCalc suck. 



  
```

Mauris placerat in velit a pharetra. Nulla viverra bibendum lectus, ut lobortis nibh. Aenean malesuada, orci vitae dictum placerat, enim justo aliquet leo, ac tincidunt augue ipsum et mi. Nullam in pharetra metus, nec molestie felis. Duis varius rhoncus pulvinar. Integer semper urna sed tellus rhoncus, vel suscipit odio iaculis. Mauris iaculis neque pellentesque ex convallis venenatis. Sed ultrices euismod fermentum. Ut feugiat risus at turpis tincidunt viverra.


```{r}
# Create a barplot of the number of fish encounters by species
ggplot(data = fish_encounters, aes(x = station, y = seen)) +
  geom_bar(stat = "identity") +
  labs(x = "Species", y = "Number of Fish Encounters")
  
```

Donec sagittis suscipit orci in ultrices. Vivamus iaculis eleifend est et dictum. Fusce non urna at lacus dapibus tincidunt ac sed nibh. Donec luctus lectus elit, ac ultrices tortor mollis sit amet. Nulla finibus lacinia arcu. Proin sollicitudin non odio et pulvinar. Maecenas finibus faucibus finibus. Ut elementum metus velit, a sodales sapien commodo et. Vivamus pulvinar, sapien et porta cursus, elit est luctus nisi, eget fringilla odio diam et enim. Sed sodales suscipit bibendum. Morbi ut metus semper, laoreet felis sodales, consequat dolor.
