---
title: "Portfolio 1: Math-ing in R"
---

>  In this portfolio piece, I will demonstrate how to do some math in R that I find particularly enjoyable. It turns out it is NOT enjoyable in R. 

Let's load the relevant packages first, among these are the package mosaiCalc which will do some of the heavy lifting.

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidyr)
library(mosaicCalc)
library(pracma) #for RREF 
#this is also helpful and necessary 
help(package="mosaicCalc")
library(knitr)
library(kableExtra)
library(rgl)
library(matlib)
```

lets slap together a system of equations: -2x + y - z = 4; x + 2y + 3z = 13; 3x + z = -1 and make a matrix 
```{r}

A <- matrix(c(-2,1,3,
              1,2,0,
              -1,3,1,
              4,13,-1), nrow=3)


```

To solve this system, we just take the Reduced Row Echelon Form of the matrix (RREF). We get exactly one solution, x=-1, y=4, z=2
```{r}
kable(rref(A))
```

Lets do it with one more system: x-2y+3z = -2; -x+y-2z=3, 2x-y+3z = -7
```{r}
B <- matrix(c(1,-1,2,
              -2,1,-1,
              3,-2,3,
              -2,3,-7), nrow=3)

rref(B)


```
This is a dependent system with infinitely many solutions. let z = t, then x = -t-4, y = t-1, and we get solutions for any real number t. 

Now I will try to plot this stuff, starting with the first example: 

```{r}
#1
A <- matrix(c(-2,1,3,
              1,2,0,
              -1,3,1), nrow=3)

b <- c( 4,13,-1)
Solve(A, b, fractions=TRUE)   

plotEqn3d(A,b, xlim=c(-1,4), ylim=c(-1,4))


#2
C <- matrix(c(1,-1,2,
              -2,1,-1,
              3,-2,3), nrow=3)
d <-c(-2,3,-7)

# Plot the system

plotEqn3d(C, d)

  
```
As it turns out, this particular method will not handle three variables, but apparently one strategy is to use heatmaps to visualize the third variable. That is something to play around with in the future. 

I'm bored of this, so I am going to practice taking some plain old boring integrals. First thing to note is that ~ is the enemy. They show up under lots of different conditions. At least in the context of differentiation, our friend "~" means "with respect to," which is fine. 

```{r}
#Tutorial Example, need to come up with my own stuff - I will just run through pauls online notes problems in here and time if I can do it faster by hand vs on R 
library(pracma) #for the integrate function, because the functions in mosiacCalc suck. 



  
```

```{r}
A <- matrix(c(2, 1, -1,
             -3, -1, 2,
             -2,  1, 2), 3, 3, byrow=TRUE)
colnames(A) <- paste0('x', 1:3)
b <- c(8, -11, -3)
showEqn(A, b)

c( R(A), R(cbind(A,b)) )          # show ranks
all.equal( R(A), R(cbind(A,b)) )  # consistent?
solve(A, b)
inv(A) %*% b
plotEqn3d(A,b, xlim=c(0,4), ylim=c(0,4))



```


```{r}
# Create a barplot of the number of fish encounters by species
ggplot(data = fish_encounters, aes(x = station, y = seen)) +
  geom_bar(stat = "identity") +
  labs(x = "Species", y = "Number of Fish Encounters")
  
```


