---
title: "Correlation Matrix"
---

I am the PSY211 TA, and one of my main responsibilities is grading research reports. I help my students use SPSS to analyse their hypotheses. In this portfolio, I will use R to create scales, and then I will create a correlation matrix among many of the variables, and then make a scatter plot matrix. 

First, I am going to load the dataset that they use for their projects: 
```{r}
library(haven)
OnlinePsychologySurveyData <- read_sav("~/Documents/TA /OnlinePsychologySurveyData.sav")
View(OnlinePsychologySurveyData)
```

Next, I will create scales for self-control, self-esteem, fear of negative evaluation, religiosity, political orientation, child-oriented, conscientousness, extraversion, emotional stability, satisfaction with life, and group affiliation seeking. 

Starting with self-control: 
```{r}
#Recode Variables 

#Reliability Analysis 

#Make the composite 
```

Next, self-esteem: 
```{r}


```

Fear of negative evaluation:
```{r}


```

Now that everything is clean and ready to go, I will produce the correlation matrix:
```{r}

```
