---
title: "Automating the Hermine Interval ACE procedure"
---

Loading packages and sourcing my BG helpers 
```{r message=FALSE}
library(BGmisc)
library(ggplot2)
library(ggpedigree)
library(OpenMx)
library(EasyMx)
library(ACEsimFit)
library(tidyr)
library(dplyr)
library(purrr)
library(polycor)
library(lattice)

source("/Users/caileyfay/Documents/Research/Github/risky_gambling/helpers.R")
source("/Users/caileyfay/Documents/Research/Github/risky_gambling/EasyMx/R/emxBehaviorGenetics.R")
source("/Users/caileyfay/Documents/Research/Github/risky_gambling/data/miFunctions.R")
```

First I create some simulated data to work on 
```{r} 
sim <- Sim_Fit(
  GroupNames = c("KinPair1", "KinPair2"),
  GroupSizes = c(2000, 2000),
  nIter = 10,
  SSeed = 2,
  GroupRel = c(.5, 0.25),
  GroupR_c = c(1, 1),
  mu = c(15, 15), #can I assume that the means for both groups are the same, even though they are slightly different in reality 
  ace1 = c(.3, .2, .5),
  ace2 = c(.3, .2, .5),
  ifComb = FALSE,
  lbound = FALSE,
  saveRaw = TRUE
)

# Convert the results element to a dataframe
# sim_df1 <- as.data.frame(sim$Iteration1$data)
# sim_df2 <- as.data.frame(sim$Iteration2$data)
# sim_df3 <- as.data.frame(sim$Iteration3$data)
# sim_df4 <- as.data.frame(sim$Iteration4$data)
# sim_df5 <- as.data.frame(sim$Iteration5$data)
# sim_df6 <- as.data.frame(sim$Iteration6$data)
# sim_df7 <- as.data.frame(sim$Iteration7$data)
# sim_df8 <- as.data.frame(sim$Iteration8$data)
# sim_df9 <- as.data.frame(sim$Iteration9$data)
# sim_df10 <- as.data.frame(sim$Iteration10$data)

#figuring out how to do it more efficiently, less copy paste nonesense 
#nah this is annoying 
#sim is a list. I want it to do something so its for every iteration in the sim list, create a data frame and label it sim_df1 .... sim_dfi 
#ok turns out its not that hard with the map function 

df_list <- map(sim, ~ as.data.frame(.x$data))
#list2env(df_list, envir = .GlobalEnv)


```

The next step is to make it plausible for drug initation data - that means some people need to have NAs to represent non-use. I also have to make this into a loop or something more efficient. The goal is to scale it up so that when I do the bigger simulations for the BGA poster I have tricks to save myself work 
```{r}
for (i in 1:length(df_list)) {
  
  df_list[[i]] <- df_list[[i]] %>%

 mutate(y1 = case_when(
    R == .5 ~ ifelse(runif(n()) < 0.20, NA, y1), 
     R == .25 ~ ifelse(runif(n()) < 0.10, NA, y1)))  %>%
    
  mutate(y2 = case_when(
    is.na(y1) ~ NA, 
    TRUE ~ y2
  )) %>%
  mutate(Ord_1 = case_when(
    y1 <= 14 ~ 1,
  y1 <= 15 ~ 2,
  y1 <= 16 ~ 3,
  y1 >= 17 ~ 4,
  is.na(y1) ~ 4,
  ),
  Ord_2 = case_when(
    y2 <= 14 ~ 1,
  y2 <= 15 ~ 2,
  y2 <= 16 ~ 3,
  y2 >= 17 ~ 4,
 is.na(y2) ~ 4,
  )) %>%
  mutate(RCoef = R)
}

list2env(df_list, envir = .GlobalEnv)
```

Traditional way of getting an Ordinal Ace from sim_df
```{r}
sim_df <- Iteration1 #relic of past naming convention but too much work to change across files at this point 
source("/Users/caileyfay/Documents/Research/Github/risky_gambling/sim_df_ordinal_ACE.r")
sumACE
table(sim_df$Ord_1)
table(sim_df$Ord_2)
histogram(sim_df$Ord_1)
histogram(sim_df$y1)
```

Interval Ace 
```{r}
source("~/Documents/Research/Github/risky_gambling/sim_df_interval_ACE.r")
sumACE
```

ACE with easymx - emxTwinModel function already modified 
```{r}
model <- emxTwinModel(model="Cholesky", relatedness ='RCoef', data=Iteration1, run=TRUE, use=c('y1','y2'), name='model', components='ACE')
summary(model)
```

Now I want to repeat this for the other 9 iterations 
```{r}
run_int <- function(df) {
 
   model <- emxTwinModel(model="Cholesky", relatedness ='RCoef', data=df, run=TRUE, use=c('y1','y2'), name='model', components='ACE')
   
}

results <- lapply(df_list, run_int)

#now I have a big list of model outputs, but I can't do anything with it. I want means, parameters 
R1 <- summary(results$Iteration1)
estimates_df1 <- R1$algebra

#now scale it up so I don't have to repeat manually ten times 
#can't figure it out yet with lapply so I am going to for loop it 

estimates_list <- lapply(results, function(fit) {
  s <- summary(fit)
  return(s$parameters) 
})

#fuck it 
R1 <- summary(results$Iteration1)
estimates_df1 <- R1$parameters
R2 <- summary(results$Iteration2)
estimates_df2 <- R2$parameters
R3 <- summary(results$Iteration3)
estimates_df3 <- R3$parameters
R4 <- summary(results$Iteration4)
estimates_df4 <- R4$parameters
R5 <- summary(results$Iteration5)
estimates_df5 <- R5$parameters
R6 <- summary(results$Iteration6)
estimates_df6 <- R6$parameters
R7 <- summary(results$Iteration7)
estimates_df7 <- R7$parameters
R8 <- summary(results$Iteration8)
estimates_df8 <- R8$parameters
R9 <- summary(results$Iteration9)
estimates_df9 <- R9$parameters
R10 <- summary(results$Iteration10)
estimates_df10 <- R10$parameters

#just in case I want to do something with this 
eh <- rbind(estimates_df1,estimates_df2, estimates_df3, estimates_df4, estimates_df5, estimates_df6, estimates_df7, estimates_df8, estimates_df9, estimates_df10)
eh <- eh %>%
  mutate(It = rep(1:10, each = 4))

A <- eh %>%
  filter(name == "sqrtA11") #these need to be transformed so that they are actually the A ( or C or E ) so I atually probably need to approach this diffrently 

C <-eh %>%
  filter(name == "sqrtC11")

E <- eh %>%
  filter(name == "sqrtE11")
```


now that the parameters are workable, I want to graph them as 3 overlapping density distributions 
```{r}
ggplot(A, mapping = aes(x=Estimate)) +
  geom_density() +
  xlim(0, 1)

ggplot(C, mapping = aes(x=Estimate)) +
  geom_density() +
  xlim(0, 1)

ggplot(E, mapping = aes(x=Estimate)) +
  geom_density() +
  xlim(0, 1)

#Dubious 
ggplot(eh, mapping = aes(x = Estimate, fill = name)) +
  geom_density() +
  xlim(0,1) +
  scale_fill_brewer() +
  theme_minimal()

```


Need to do the ordinal with easyMx - will need to modify the function 
```{r}

```

Trying out this function 
```{r}



```

```{r}



```

