---
title: "Tidy Tuesday: Netflix"
---

For this portfolio piece, I am handing over the "choose your own adventure" reins to the Tidy Tuesday Repo, because I hate making decisions.

Here are some of the things TidyTuesday wanted me to think about:

-   [ ] What is the relation between time since release and current viewing performance? Do older shows have staying power or do newer releases dominate?
-   [ ] How does the timing of a show's release (month/season) relate to engagement metrics?
-   [ ] How does the performance of globally available content compare to regionally restricted titles? Do global releases consistently outperform regional ones?
-   [ ] How do the top movie performers in each reporting period compare? Are the same kinds of movies consistently popular?
-   [ ] How does viewing change across seasons for shows that appear in multiple Netflix engagement reports? Do later seasons of popular shows maintain, gain, or lose audience engagement compared to their earlier seasons?

First I will read in the data from Github:

```{r message=FALSE}
movies <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/movies.csv')
shows <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-07-29/shows.csv')

library(ggplot2)
library(polycor)
library(tidyr)
library(dplyr)
library(tidytuesdayR)
library(stringr)
library(stringi)
library(ggthemes)
```

Next I am going to get a sense of the lay of the land. From the following codes, I learned:

-   [] We have data reports every 6 months on shows, from July of 2023 to June 2025. This means I can evaluate changes in engagement chunked into 6 month periods.
-   [] The hours_viewed numbers are crazy, so adjusting them somehow will be helpful. Some options to consider would be making a proportion of hours viewed by runtime, or number of views?
-   [] Runtime is a character, and thats a problem I may or may not choose to address

```{r}
table(shows$report)
summary(shows$hours_viewed)
str(shows$runtime)

#adjusting the hours viewed into something more manageble 
shows$end_date <- now()
shows <- shows %>%
  mutate(prop_hrs_viewed = hours_viewed / views,
         release_year = year(release_date),
         release_month = month(release_date),
         age = as.duration(interval(release_date, end_date)))

shows <- shows %>%
  mutate(weeks_old = as.numeric(age / 604800))

#filter for love is blind 
LOB <- shows %>%
  filter(stri_startswith_fixed(title, "Love is Blind: Season", case_insensitive = TRUE))

#Filter for too hot too handle 
Hot <- shows %>%
  filter(str_starts(title, "Too Hot to Handle"))

Dating_shows <- bind_rows(Hot, LOB)

Dating_shows <- Dating_shows %>%
  mutate(category = case_when(
 str_starts(title,"Too Hot to Handle") ~ 1,
  str_starts(title,"Love Is Blind") ~ 2,
    .default = 3
  ))

#we can see about release date becoming numeric by using date since origin? 

```

There are duplicates with different report dates. We can use this to see how viewership changes for each season as it gets older: 
```{r}
LOB4 <- LOB %>%
  filter(stri_startswith_fixed(title, "Love is Blind: Season 4", case_insensitive = TRUE))

ggplot(data = LOB4, mapping = aes(x=report, y=views)) +
geom_col(fill = "pink", color = "black") +
  labs(title = "Love Is Blind: Season 4 Views Since Release",
         x = "Netflix Report Date",
         y= "Number of Views") 
```

Now I want to do this so its every season displayed like this and faceted 
```{r}
#I think I need to make a variable that selects the four reports for each season 
LOB <- LOB %>%
  mutate(Season = case_when(
    title == "Love Is Blind: Season 1" ~ 1,
    title == "Love Is Blind: Season 2" ~ 2,
    title == "Love Is Blind: Season 3" ~ 3,
    title == "Love Is Blind: Season 4" ~ 4,
    title == "Love Is Blind: Season 5" ~ 5,
    title == "Love Is Blind: Season 6" ~ 6,
    title == "Love Is Blind: Season 7" ~ 7,
    title == "Love Is Blind: Season 8" ~ 8,
    TRUE ~ NA_real_
  ))

ggplot(data = LOB, mapping = aes(x=report, y=prop_hrs_viewed)) +
geom_col(fill = "pink", color = "black") +
  facet_wrap(~ Season)
  labs(title = "Love Is Blind: Season 4 Views Since Release",
         x = "Netflix Report Date",
         y= "Hours per view") 

```

Which is more popular: Love is Blind or Too Hot Too Handle?
```{r}
ggplot(LOB, mapping = aes(x=release_date, y=prop_hrs_viewed)) +
  geom_bar()
```

Release Timing and Engagement - [ ] I am going to first look at the past year, filtered by month. - [ ] Depending on how that goes, I am going to look at it averaged over years (i.e. average engagement of shows released in December across however many years of data I have)

```{r}
#release month - bubble scatter plot where the size of the bubbles is the release year 
# x can be release month 
# y can be the views 

# thank you https://r-graph-gallery.com/320-the-basis-of-bubble-plot.html
shows %>%
  ggplot(aes(x=release_year, 
             y=views,
             size = release_month,
             color = report)) +
  geom_point(alpha = .3) +
  scale_size(range = c(1,10), name = "Release Month")  +
  labs(title = "Release Timing and Engagement",
       x = "Release Year",
       y = "Views")

shows %>%
  ggplot(aes(x=release_month, y=hours_viewed, size=re, fill=available_globally)) +
    geom_point(alpha=0.5, shape=21, color="black") +
    scale_size_area(name="Year") +
    theme_wsj() +
    theme(legend.position="bottom") +
    ylab("Hours Viewed") +
    xlab("Month of Release") +
    theme(legend.position = "bottom")
```

What is the relation between time since release and current viewing performance? Do older shows have staying power or do newer releases dominate?
```{r}
#lets use the weeks_old variable for age 
#to get current viewing performance I need to filter for the most recent report, which was 2025Jan-Jun
#then, I think we start with a simple scatter of the show and its views on the y and months old on the x, and build from there 

shows %>%
  filter(report == "2025Jan-Jun",
         views >= 10000000,
         views <= 20000000) %>%
  ggplot(aes(x=weeks_old, y=views)) +
  geom_point(alpha = .7) +
   labs(title = "Among the most popular shows, does age relate to views in 2025?",
       x = "Weeks Old",
       y = "Views")


#doesn't show much of a relation, but also there is so much data. might help to color dots by category 

Dating_shows %>%
  filter(report == "2025Jan-Jun",
         views >= 3000000) %>%
  ggplot(aes(x=weeks_old, 
             y=views,
             color = category)) +
  geom_point(alpha = .7) +
  labs(title = "Do Love is Blind and Too Hot Too Handle Have Any Staying Power?",
       x = "Weeks Old",
       y = "Views")
#there is not different staying power for old vs new shows when comparing the love is blind series and too hot too handle 

#newer seems to dominate 

```