---
title: "Portfolio 4: Using Shiny App to Make an Interactive Dashboard of Some ESM Personality Data"
---
I was once again inspired by something we talked about in Will's Personality Seminar. SO, for this portfolio, I will be doing an abrieviated version of the density distribution approach for mapping the relationship between trait standing and trait manifestation. The data is coming from my favorite guinea pig, my boyfriend, who was nice enough to fill out some short pseudo personality state assessment that I made. Here are my learning objectives for this portfolio piece...lets see if I succeed! 

- [ ] I want to learn how to make an interactive dashboard with the Shiny App. This is how I will visualize mr. boyfriends data. 
- [ ] I also wanted to learn how to do a survey in such a way that it can be imported straight into r as it updates, and it turns out it was shockingly easy! Spoiler alert: I consider myself a qualtrics girly but google forms and some random post on stackoverflow did me a real solid. 
- [ ] I wanted to test out the density distribution method on a real person I actually know. I am aware it's not empirical, but I thought it would be fun. 
- [ ] My last learning objective is figuring out how to generate lots of data points based on a smaller number of real, existing datapoints. Me and mr. boyfriend filled out the survey around 5-10 times each, and from that I generated 1000 datapoints for each of the big 5 traits. 

To get started with the Shiny app, I used this tutorial https://www.youtube.com/watch?v=VTI84Pqk3SI&t=1s by Grace Thompkins that was super helpful. 

First, of course, I will load up my packages like always. In a separate file, I created the link between my google form responses and then cleaned it up a little bit. this is the document I am sourcing. It just makes life easier for when I am trying to set up the app. 
```{r}
library(shiny)
library(googlesheets4)
library(janitor)
library(dplyr)
library(tidyr)
library(gsheet)
library(ggplot2)
library(synthpop)

source("/Users/caileyfay/Documents/LabsD4P/Data-Science-for-Psych-Portfolio/P4_ESM/cleaning_ESM.r")

```



```{r}
#playing around with the bones of the graph I will create

#I want to be able to click to see the profile of mine vs ethans, seperately 
#I want a graph of mine mapped over his for one of the traits, or a faceted chart that shows ours stacked on top of each other for each of the five traits 
#I want people to be able to select the personalty characteristic of interest 
#the density thing is going to be a histogram for a trait
#I also want plots of the time of day vs the level of the trait 
#to help me get a concrete sense before I code something more generalized in the interactive graph: 
#initial plots with the real data
ESM %>%
  filter(who_are_you =="Cailey") %>%
ggplot(aes(x=hour, y=Extraversion)) +
  geom_point() +
  geom_line() #making a scatter plot of the time of day and level of trait 

ESM %>%
  filter(who_are_you =="Cailey") %>%
ggplot(aes(x=Extraversion)) +
  geom_histogram() #density distribution for one of the traits? 

ESM %>%
  ggplot(mapping = aes(x=Extraversion),
       color = who_are_you,
       fill = who_are_you) +
  geom_density(alpha = 0.7) #comparing Cailey and Ethan density distributions for a given trait this doesn't work rn 
#I would like to figure out how to facet it so that we get five of these 
```
```{r}
#generating synthetic data so I have more points to work with 
#trying out snythpop from: https://www.synthpop.org.uk/get-started.html
synthetic_dataset <- syn(ESM, k=1000)
write.syn(synthetic_dataset, filename = "my_synthetic_data", filetype = "csv")
 syn_data <- read_csv("my_synthetic_data.csv")
```
```{r}
#revisualizing 
 syn_data %>%
  filter(who_are_you =="Cailey") %>%
ggplot(aes(x=hour, y=Extraversion)) +
  geom_point()  #making a scatter plot of the time of day and level of trait 

#Histograms 
syn_data %>%
  filter(who_are_you =="Cailey") %>%
ggplot(aes(x=Extraversion)) +
  geom_histogram() +
  labs(title = "Cailey's Extraversion Distribution")#density distribution for one of the traits? 

syn_data %>%
  filter(who_are_you =="Ethan") %>%
ggplot(aes(x=Extraversion)) +
  geom_histogram()  +
  labs(title = "Ethan's Extraversion Distribution")

#Overlapping Density Distributions For Each Trait 
syn_data %>%
  ggplot(mapping = aes(x=Extraversion, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Extraversion Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal()

syn_data %>%
  ggplot(mapping = aes(x=Agreeableness, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Agreeableness Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal()

syn_data %>%
  ggplot(mapping = aes(x=Conscientiousness, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Conscientiousness Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal()

syn_data %>%
  ggplot(mapping = aes(x=Emotional_Stability, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Emotional Stability Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal()

syn_data %>%
  ggplot(mapping = aes(x=Openness, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Openness to New Experiences Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal()
#I also want to do these so you get the density of all of the traits but for one person on the graph - probably needs to happen with faceting but we shall see 

```

