---
title: "Portfolio 4: Using Shiny App to Make an Interactive Dashboard of Some ESM Personality Data"
runtime: shiny
---
I was once again inspired by something we talked about in Will's Personality Seminar. SO, for this portfolio, I will be doing an abrieviated version of the density distribution approach for mapping the relationship between trait standing and trait manifestation. The data is coming from my favorite guinea pig, my boyfriend, who was nice enough to fill out some short pseudo personality state assessment that I made. Here are my learning objectives for this portfolio piece...lets see if I succeed! 

- [ ] I want to learn how to make an interactive dashboard with the Shiny App. This is how I will visualize mr. boyfriends data. 
- [ ] I also wanted to learn how to do a survey in such a way that it can be imported straight into r as it updates, and it turns out it was shockingly easy! Spoiler alert: I consider myself a qualtrics girly but google forms and some random post on stackoverflow did me a real solid. 
- [ ] I wanted to test out the density distribution method on a real person I actually know. I am aware it's not empirical, but I thought it would be fun. 
- [ ] My last learning objective is figuring out how to generate lots of data points based on a smaller number of real, existing datapoints. Me and mr. boyfriend filled out the survey around 5-10 times each, and from that I generated 1000 datapoints for each of the big 5 traits. 

To get started with the Shiny app, I used this tutorial https://www.youtube.com/watch?v=VTI84Pqk3SI&t=1s by Grace Thompkins that was super helpful. 

First, of course, I will load up my packages like always. In a separate file, I created the link between my google form responses and then cleaned up the data a little bit and sourced it. 

```{r message=FALSE, warning=FALSE, results='hide'}
library(shiny)
library(shinydashboard)
library(googlesheets4) #in love with this
library(janitor)
library(dplyr)
library(tidyr)
library(gsheet)
library(ggplot2)
library(synthpop) #also in love with this - very cool discovery 
library(flexdashboard)
library(haven)

source("/Users/caileyfay/Documents/LabsD4P/Data-Science-for-Psych-Portfolio/P4_ESM/cleaning_ESM.r")
```

Next I generated lots of data using the synthpop package, which I learned about here:

https://www.synthpop.org.uk/get-started.html

You can find the code in the cleaning_ESM.r file. 

I made some fun charts to see how my synthetic data turned out. I start out super simple. 


```{r echo=FALSE, message=FALSE, warning=FALSE}
#line is weird with this data, but lets try bar for extraversion and time of day 

syn_data %>%
  ggplot(mapping = aes(x=hour, y=Extraversion, fill = who_are_you)) +
  geom_col() + 
  scale_fill_manual(values =c("lightblue","#2A52BE")) +
  labs(title = "Extraversion at Differnt Times of Day",
       x = "Hour",
       y= "Extraversion",
       fill = "Person") +
  theme_minimal()

#Histograms 
syn_data %>%
  filter(who_are_you =="Cailey") %>%
ggplot(aes(x=Extraversion)) +
  geom_histogram(fill="lightblue") +
  theme_minimal() +
  labs(title = "Cailey's Extraversion Distribution",
       x="Extraversion Level",
       y="Frequency")

syn_data %>%
  filter(who_are_you =="Ethan") %>%
ggplot(aes(x=Extraversion)) +
  geom_histogram(fill="#2A52BE")  +
  theme_minimal() + 
  labs(title = "Ethan's Extraversion Distribution",
        x="Extraversion Level",
       y="Frequency")

#Overlapping Density Distributions For Each Trait 
syn_data %>%
  ggplot(mapping = aes(x=Extraversion, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Extraversion Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal() +
  scale_fill_manual(values =c("lightblue","#2A52BE"))

syn_data %>%
  ggplot(mapping = aes(x=Agreeableness, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Agreeableness Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal() +
   scale_fill_manual(values =c("lightblue","#2A52BE"))

syn_data %>%
  ggplot(mapping = aes(x=Conscientiousness, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Conscientiousness Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal() +
   scale_fill_manual(values =c("lightblue","#2A52BE"))

syn_data %>%
  ggplot(mapping = aes(x=Emotional_Stability, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Emotional Stability Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal() +
   scale_fill_manual(values =c("lightblue","#2A52BE"))

syn_data %>%
  ggplot(mapping = aes(x=Openness, fill= who_are_you)) +
  geom_density(alpha = 0.5) +
  labs(title = "Ethan and Cailey's Openness to New Experiences Density Distributions",
       y = "Density",
       fill = "Subject") +
  theme_minimal() +
   scale_fill_manual(values =c("lightblue","#2A52BE"))
#I also want to do these so you get the density of all of the traits but for one person on the graph - probably needs to happen with faceting but we shall see 

```

Nothing about the way I conducted this was foolproof, but this little snapshot makes it seem like we have fairly similar personality profiles. 

Finally, I made an interactive dashboard. Its not very complex, but bonus points for it taking me forever to figure out??? 

### GAHH I can't see to get rid of this ugly warning message 

Point to a different repo that hosts the shiny app and it will do better? 

```{r echo=FALSE, message=FALSE, warning=FALSE}
 suppressWarnings(
   suppressMessages(shiny::shinyAppFile("P4_ESM/app.R")))

```

