---
title: "Factor analysis with my NLSY variables"
#output: 
#  html_document:
 # theme: lumen  # Options: cerulean, flatly, spacelab, yeti, etc.
 #   highlight: tango # Options: pygments, kate, zenburn, etc.
 
 #circle back to figure out how to do a fun theme 
---

So far I've done it w lavaan, but I will also go back and try it with OpenMx

 I will be following this tutorial to learn how to do factor analysis in R: https://bookdown.org/sz_psyc490/r4psychometics/factor-analysis.html but of course I will be applying this to my own data. I want to know about the factor structure of the kid and parent risk variables that I informally created scales for (and then used in some of my ACE models - scary). Probably helpful to know about the potential consequences of those decisions, even though the Drug initiation stuff that I am currently using isn't subject to the same concerns. 

First I need to get my NLSY data loaded into here. Its already been cleaned and prepped in these scripts.  
```{r}
library(dplyr)
library(psych)
library(lavaan)
library(corrplot)
library(semPlot)
library(ggplot2)

source('~/Documents/Research/GitHub/risky_gambling/data/CBRandYBR/CBRandYBR.R')
source('~/Documents/Research/GitHub/risky_gambling/data/GeneralRisk79/GeneralRisk79.R')
```

I am first going to do an EFA with the kid risk variables for one year - lets take 2010. This probably isnt enough items.  
```{r}

kidFA_2010 <- df_CRBandYRB %>% select(c("CRB1_2010","CRB2_2010","CRB3_2010","CRB4_2010","CRB5_2010", "CRB6_2010"))

#efa_1 <- factanal(x = kidFA_2010, factors = 1) this doesn't work 
```

This does not run because I have a lot of missingness, and this particular function I am using doesn't like it. I'm going to be brave and do na.omit and see what happens. Good news! I still have 428 observations. I'm fine with these consequences. 
```{r}
bye <- na.omit(kidFA_2010)
efa_1 <- factanal(x = bye, factors = 1)
efa_1
```
These results suggest that the one factor structure is OK, which is nice. CRB1 and CRB3 aren't so good but I am ok with this. Next, I repeat with the parent risk. Does this also have a one-factor structure? 

```{r}
parentFA_2010 <- df_GeneralRisk79 %>% select(c("RDriving_2010","RFinancial_2010","ROccupation_2010","RHealth_2010","RTrust_2010","RRomantic_2010","RLifeChanges_2010","RBets_2010"))

bye_2 <- na.omit(parentFA_2010) #this one has 6,988 still, which is nice. 

efa_2 <- factanal(x = bye_2, factors = 1)
efa_2

efa_21 <- factanal(x = bye_2, factors = 2)
efa_21

```
In a shocking turn of events, the one factor structure also looks fine. The two structure does not. Cool. At least among these few types of risk taking items, they are maybe capturing the same thing. I wish that the kids and parents answered the same items, so you could see if the structure differs by age cohort - is the nature of different types of risk taking something that varies by age group? 

--------------------------------------------------------------------------------------------------------------

I learned you can use the Lavaan package to do this with SEM, which is another way to get around the missingness issue, but it wants me to propose a structure, which I didn't have before. Now though, I am going to give it a go. I am using this tutorial to help me figure out how to apply it to my own variables:  https://lavaan.ugent.be/tutorial/efa.html

```{r}
#not SEM
efa.model <- 'efa("efa")*f1 =~ RDriving_2010 + RFinancial_2010 + ROccupation_2010 + RHealth_2010 + RTrust_2010 + RRomantic_2010 + RLifeChanges_2010 + RBets_2010'

fit <- cfa(efa.model, data = parentFA_2010)
summary(fit, standardized = TRUE)
```
The new tutorial showed me a different function to use, so the missingness is less of a problem! yey. It looks like this efa function just used the complete cases, so the sample ends up being the same as the na.omit route, but I don't have to to the deletion, which I like. 

Its also interesting to see that the factor loadings are strong (st.all I think) but the model fit is significant, which is no bueno in this context? Heres the scree plot: 

```{r}
scree(parentFA_2010)
scree(kidFA_2010)
```
This is a one factor structure all right - which my scale is only 8 items so its not that surprising - not that many possible options. However, risk taking in romantic relationships versus driving versus gambling seem like they could be different. I am entertained by this. 
--------------------------------------------------------------------------------------------------------------

Now that I have it sort of figured it out, I am going to add in a little more complexity. I want to take individuals scores on the variables when they were 12, and those are the scale years that will go into the FA. Not sure if its possible but it would make the most sense to be doing the analysis in such a way that ages are a little more consistent 

Revisting this later - I like the parent risk vars much better and it helps that there are more of them. 
```{#r}
kidFA <- df_CRBandYRB %>% select(-c("CASEID","MCASEID","Race","Sex","CBirthYear","TargetRisk","Age","Year12")) %>% select(-contains("scale")) 

#kidFA[] <- lapply(kidFA, as.numeric(kidFA))

kidFA <- na.omit(kidFA)

factanal(x = kidFA, factors = 1)
```

Or can I just put everything in and see what happens? 

Turns out there is a thing called Longitudinal CFA, which asks: is the scale measuring the same thing across time? This tutorial looks promising for learning how to do it, so that's where I will start. Note: I am modeling my analysis after the steps they outline, but I am applying it to my own data. Hope this is an ok strategy and not cheating: 
https://quantdev.ssri.psu.edu/sites/qdev/files/LongitudinalMeasurementInvariance_2017_1108.html

There are three different years of general risk variables, and I am going to select these and make a new frame. 
```{r}
LCFA_vars <- df_GeneralRisk79 %>% select(-c("CASEID","Cohort","Race","Sex", "GenRisk2012","GenRisk2010","GenRisk2014","Age1B_2022"))

describe(LCFA_vars)

#interesting thing I had noticed before - we have lots of respondents for 2010 and then a sharp drop off for 2012 and 2014. I hate the NLSY. This is probably going to be a problem. 

cor(LCFA_vars, use = "pairwise.complete.obs")
```

Absolutely ridiculous. I am going to go out on a whim and say that no one was given this set of risk questions for more than one year. Not giving up yet - I have depression variables to work with 
```{r}
source('~/Documents/Research/GitHub/risky_gambling/data/depression79/depression79.R')

depLFA <- depression79 %>% select(-c("CASEID_1979"))

#describe(depLFA)
cor(depLFA, use = "pairwise.complete.obs") #this looks better at least 
#making a pretty plot because why not
corrplot(cor(depLFA, use = "pairwise.complete.obs"), order = "original", tl.col='black', tl.cex=.75)

#the sample size is feasible but this group is not very depressed at all 
#for example - this is so right skewed its not even funny
ggplot(depLFA, mapping = aes(x=Q2_2020)) + geom_histogram()


```

Now that I have found some longitudinal data that checks out, I can start modeling. Starting with a Configural Invariance Model - I gather this is looking at: do we have the same factor structure for each group (or in this case, the group is the year). Is the 2020 factor structure reasonably similar to the 2022 factor structure? 
```{r}
configural <- '
# Define the latent factors.
#the lamba and the NA stuff is saying not to fix the first factor loading to one. 
f1 =~ NA*Q1_2020 + lambda1*Q1_2020 + Q2_2020 + Q3_2020 + Q4_2020 + Q5_2020 + Q6_2020 + Q7_2020 + Q8_2020
f2 =~ NA*Q1_2022 + lambda1*Q1_2022 + Q2_2022 + Q3_2022 + Q4_2022 + Q5_2022 + Q6_2022 + Q7_2022 + Q8_2022

#I think there is a one factor structure for CESD and I think time1 and time2 will have the same one-factor structure, so thats what the stuff above is saying. I know this is not an exciting hypothesis by any means, its just practice 

# Intercepts
Q1_2020 ~ i1*1
Q2_2020 ~ 1
Q3_2020 ~ 1
Q4_2020 ~ 1
Q5_2020 ~ 1
Q6_2020 ~ 1
Q7_2020 ~ 1
Q8_2020 ~ 1
Q1_2022 ~ i2*1
Q2_2022 ~ 1
Q3_2022 ~ 1
Q4_2022 ~ 1
Q5_2022 ~ 1
Q6_2022 ~ 1
Q7_2022 ~ 1
Q8_2022 ~ 1

# Unique Variances #still a little confused about what this code is accomplishing 
Q1_2020 ~~ Q1_2020
Q2_2020 ~~ Q2_2020
Q3_2020 ~~ Q3_2020
Q4_2020 ~~ Q4_2020
Q5_2020 ~~ Q5_2020
Q6_2020 ~~ Q6_2020
Q7_2020 ~~ Q7_2020
Q8_2020 ~~ Q8_2020
Q1_2022 ~~ Q1_2022
Q2_2022 ~~ Q2_2022
Q3_2022 ~~ Q3_2022
Q4_2022 ~~ Q4_2022
Q5_2022 ~~ Q5_2022
Q6_2022 ~~ Q6_2022
Q7_2022 ~~ Q7_2022
Q8_2022 ~~ Q8_2022


# Latent Variable Means
f1 ~ 0*1 #factor at time 1 is fixed to zero to be a baseline 
f2 ~ 1

# Latent Variable Variances and Covariance
f1 ~~ 1*f1
f2 ~~ f2
f1 ~~ f2 #correlation between the latent var at time 1 and time 2 
'

#Model fitting
fit_configural <- cfa(configural, data = depLFA, mimic = "mplus")
summary(fit_configural, fit.measures = TRUE, standardize = TRUE)

```

It ran! I don't know what it did, so now I need to figure out what it means / accomplishes in this first step. 

New plotting technique
```{r}
#Model Diagram 
semPaths(fit_configural, what="std", 
         sizeLat = 7, sizeMan = 7, edge.label.cex = .75,
         style = "lisrel")

#trying to figure out what settings make the EV circles (I guess triangles here) proportional to how much unexplained var there is - but perhaps doesn't work w this data bc generaly EV is mostly similar 
```
Now to try to make sense of these results: I think I see
 [] the latent variable depression at time one and time two are correlated at .57. people who tend to be more depressed at time 1 are also usually more depressed at time 2. 
 [] there are similar factor loadings at time one and time two 
 [] the configural thing that I just ran is looking to see if the factor structure is the same across time 
 [] it looks like there are lots of other types (i.e. weak and strong invariance models) that can compare factor loadings, latent variable means, etc at different times 
 []its a little weird - there is plenty of info online for getting the syntax, but not for interpretting what the output means - I don't think Mike is going to be teaching us about this so I am on my own 
 [] CESD is a really solid measure, so I was expecting clean results and it looks like I got them. At this point I am not going to get into interpretting the inferential stuff and the model fit. the diagram is fairly informative for my currently level of understanding. 
 
 