---
title: "Factor analysis with my NLSY variables"
---

I will be following this tutorial to learn how to do factor analysis in R: https://bookdown.org/sz_psyc490/r4psychometics/factor-analysis.html. This tutorial uses factanala() from the built in stats package. Later in this portfolio,  I follow a tutorial that uses lavaan. If time allows, I will also try to figure it out with OpenMx. 
 
I want to know about the factor structure of the kid and parent risk variables from the NLSY. I will also try out a longitudinal CFA on depression of the NLSY79 in 2020 and 2022. 

```{r include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
```{r include=FALSE}
library(dplyr)
library(psych)
library(lavaan)
library(corrplot)
library(semPlot)
library(ggplot2)
library(broom) #to present my FA results a little cleaner 
library(knitr)
library(kableExtra)

source('~/Documents/Research/GitHub/risky_gambling/data/CBRandYBR/CBRandYBR.R')
source('~/Documents/Research/GitHub/risky_gambling/data/GeneralRisk79/GeneralRisk79.R')
```

## EFA For the Child Risk Behavior Index. 

I am first going to do an EFA with the child risk behavior index for 2010. 

I use the factanal() function to see what the one-factor structure looks like. This function requires complete cases, so I used na.ommit, which yielded 428 observations. 

```{r echo=FALSE}
kidFA_2010 <- df_CRBandYRB %>% select(c("CRB1_2010","CRB2_2010","CRB3_2010","CRB4_2010","CRB5_2010", "CRB6_2010"))
scree(kidFA_2010)
```

```{r include=FALSE}
#efa_1 <- factanal(x = kidFA_2010, factors = 1) this doesn't work 
```

The following table shows the items, variance unexplained by common factors (uniqueness), and their factor loadings (fl1).

```{r echo=FALSE}
bye <- na.omit(kidFA_2010)
efa_1 <- factanal(x = bye, factors = 1)
output <- tidy(efa_1)
kable(output)
```

These results suggest that the one factor structure isn't amazing. Next, we see how a 2 factor structure looks. It doesn't look any better. 

```{r echo=FALSE}
bye <- na.omit(kidFA_2010)
efa_15 <- factanal(x = bye, factors = 2)
output <- tidy(efa_15)
kable(output)
```

Since the parallel analysis suggested four factors, I tried to test that out, but it isn't possible with 6 items. 
```{r echo=FALSE}
#bye <- na.omit(kidFA_2010)
#efa_14 <- factanal(x = bye, factors = 4)
#kable(tidy(efa_14))
```

For the sake of time, I am not going to get into the interpretation, and I am going to focus on learning + visualzing how to conduct these. Mike is currently covering EFA and CFA right now in class. 


## EFA for the NLSY79 General Risk Variables 

There are 8 risk domains measured in the NLSY79 in 2010, 2012, and 2014. They are all measured on a 1-10 scale for behaviors such as financial, romantic, and health related risk-taking. 

I selected the 8 items from 2010 and cleared out all incomplete observations. There were 6,988 observations remaining. First I checked out the scree plot and parallel analysis. 

```{r echo=FALSE}
parentFA_2010 <- df_GeneralRisk79 %>% select(c("RDriving_2010","RFinancial_2010","ROccupation_2010","RHealth_2010","RTrust_2010","RRomantic_2010","RLifeChanges_2010","RBets_2010"))
scree(parentFA_2010)
fa.parallel(parentFA_2010)
```

Next I test the one factor structure: 
```{r echo=FALSE}
bye_2 <- na.omit(parentFA_2010) #this one has 6,988 still, which is nice. 
efa_2 <- factanal(x = bye_2, factors = 1)
kable(tidy(efa_2))
```

The two factor structure: 
```{r echo=FALSE}
efa_21 <- factanal(x = bye_2, factors = 2)
kable(tidy(efa_21))
```

The one factor structure wins. *report sig stuff* These results provide some evidence that there is a latent risk-taking variable that influences many different domains of behavior.

## CFA in Lavaan 

I learned you can use the Lavaan package to do this with SEM, which is another way to get around the missingness issue, but it wants me to propose a structure, which I didn't have before. I am using this tutorial to help me figure out how to apply it to my own variables:  https://lavaan.ugent.be/tutorial/efa.html

Before doing SEM stuff though, I tried out the cfa function on the same parent risk taking variables from 2010. I know you are supposed to do the cfa on a different sample than the efa, but in this case it's just practice, so I will go ahead and give it a whirl. 

The cfa function just used the complete cases, so the sample ends up being the same as the na.omit route, but I don't have to manually do the deletion, which I like. 

```{r echo=FALSE}
#not SEM
efa.model <- 'efa("efa")*f1 =~ RDriving_2010 + RFinancial_2010 + ROccupation_2010 + RHealth_2010 + RTrust_2010 + RRomantic_2010 + RLifeChanges_2010 + RBets_2010'

fit <- cfa(efa.model, data = parentFA_2010)
summary(fit, standardized = TRUE)
#kable(tidy(fit)) #does work as well with FA output. Think about alternate presentations. 
```


Its also interesting to see that the factor loadings are strong (st.all I think) but the model fit is significant, which is no bueno in this context? 

This is a one factor structure all right - which my scale is only 8 items so its not that surprising - not that many possible options. However, risk taking in romantic relationships versus driving versus gambling seem like they could be different. I am entertained by this. 

## Longitudinal CFA with the CESD 

Turns out there is a thing called Longitudinal CFA, which asks: is the factor structure of the scale the same across time? This tutorial looks promising for learning how to do it, so that's where I will start. Hope this is an ok strategy and not cheating: 

https://quantdev.ssri.psu.edu/sites/qdev/files/LongitudinalMeasurementInvariance_2017_1108.html

There are three different years of general risk variables, and I am going to select these and make a new frame. 
```{r echo=FALSE, results='hide'}
LCFA_vars <- df_GeneralRisk79 %>% select(-c("CASEID","Cohort","Race","Sex", "GenRisk2012","GenRisk2010","GenRisk2014","Age1B_2022"))

describe(LCFA_vars)

#interesting thing I had noticed before - we have lots of respondents for 2010 and then a sharp drop off for 2012 and 2014. I hate the NLSY. This is probably going to be a problem. 

cor(LCFA_vars, use = "pairwise.complete.obs")
corrplot(cor(LCFA_vars, use = "pairwise.complete.obs"), order = "original", tl.col='black', tl.cex=.75) +
  labs(title="Correlations Among Risk Taking Variables in 2010, 2012, and 2014",
       subtitle="No participant answered these items for all three years")
```

From this, I learned that there were no complete observations. This suggests that perhaps most people were asked in 2010, and if they didn't answer in 2010, they were asked in 2012 or 2014. 


I will pivot to NLSY79 depression variables (CESD items) which have better coverage. 

```{r echo=FALSE, results='hide'}
source('~/Documents/Research/GitHub/risky_gambling/data/depression79/depression79.R')

depLFA <- depression79 %>% select(-c("CASEID_1979"))

#describe(depLFA)
cor(depLFA, use = "pairwise.complete.obs") #this looks better at least 
#making a pretty plot because why not
corrplot(cor(depLFA, use = "pairwise.complete.obs"), order = "original", tl.col='black', tl.cex=.75)

#the sample size is feasible but this group is not very depressed at all 
#for example - this is so right skewed its not even funny
ggplot(depLFA, mapping = aes(x=Q3_2020)) + geom_histogram() + theme_classic() + 
  labs(title="The NLSY is Not Very Depressed", 
       subtitle = "Example: Responses to 'I felt that I couldn't shake off the blues even with help from my family and friends'",
       x="Statements Rated from 'Rarely(0)' to 'Most of the Time(3)'",
       y="Frequency")
```


Now that I have found some longitudinal data that checks out, I can start modeling. Starting with a Configural Invariance Model - I gather this is looking at: do we have the same factor structure for each group (or in this case, the group is the year). Is the 2020 factor structure reasonably similar to the 2022 factor structure? 

```{r echo=FALSE}
configural <- '
# Define the latent factors.
#the lamba and the NA stuff is saying not to fix the first factor loading to one. 
f1 =~ NA*Q1_2020 + lambda1*Q1_2020 + Q2_2020 + Q3_2020 + Q4_2020 + Q5_2020 + Q6_2020 + Q7_2020 + Q8_2020
f2 =~ NA*Q1_2022 + lambda1*Q1_2022 + Q2_2022 + Q3_2022 + Q4_2022 + Q5_2022 + Q6_2022 + Q7_2022 + Q8_2022

#I think there is a one factor structure for CESD and I think time1 and time2 will have the same one-factor structure, so thats what the stuff above is saying. I know this is not an exciting hypothesis by any means, its just practice 

# Intercepts
Q1_2020 ~ i1*1
Q2_2020 ~ 1
Q3_2020 ~ 1
Q4_2020 ~ 1
Q5_2020 ~ 1
Q6_2020 ~ 1
Q7_2020 ~ 1
Q8_2020 ~ 1
Q1_2022 ~ i2*1
Q2_2022 ~ 1
Q3_2022 ~ 1
Q4_2022 ~ 1
Q5_2022 ~ 1
Q6_2022 ~ 1
Q7_2022 ~ 1
Q8_2022 ~ 1

# Unique Variances #still a little confused about what this code is accomplishing 
Q1_2020 ~~ Q1_2020
Q2_2020 ~~ Q2_2020
Q3_2020 ~~ Q3_2020
Q4_2020 ~~ Q4_2020
Q5_2020 ~~ Q5_2020
Q6_2020 ~~ Q6_2020
Q7_2020 ~~ Q7_2020
Q8_2020 ~~ Q8_2020
Q1_2022 ~~ Q1_2022
Q2_2022 ~~ Q2_2022
Q3_2022 ~~ Q3_2022
Q4_2022 ~~ Q4_2022
Q5_2022 ~~ Q5_2022
Q6_2022 ~~ Q6_2022
Q7_2022 ~~ Q7_2022
Q8_2022 ~~ Q8_2022


# Latent Variable Means
f1 ~ 0*1 #factor at time 1 is fixed to zero to be a baseline 
f2 ~ 1

# Latent Variable Variances and Covariance
f1 ~~ 1*f1
f2 ~~ f2
f1 ~~ f2 #correlation between the latent var at time 1 and time 2 
'

#Model fitting
fit_configural <- cfa(configural, data = depLFA, mimic = "mplus")
summary(fit_configural, fit.measures = TRUE, standardize = TRUE)
```

It ran! I don't know what it did, so now I need to figure out what it means / accomplishes in this first step. 

Finally, I tried out a new plotting technique, semPaths. 
```{r echo=FALSE}
#Model Diagram 
semPaths(fit_configural, what="std", 
         sizeLat = 7, sizeMan = 7, edge.label.cex = .75,
         style = "lisrel")

#trying to figure out what settings make the EV circles (I guess triangles here) proportional to how much unexplained var there is - but perhaps doesn't work w this data bc generaly EV is mostly similar 
```

The correlation between the latent variable depression at time 1 and time 2 is .66, and they appear to have the same factor structure. This makes sense because the CESD is a really good measure of depression, and time 1 and time 2 are only two years apart. 
```{r echo=FALSE} 
# Now to try to make sense of these results: I think I see
#  [] the latent variable depression at time one and time two are correlated at .57. people who tend to be more depressed at time 1 are also usually more depressed at time 2. 
#  [] there are similar factor loadings at time one and time two 
#  [] the configural thing that I just ran is looking to see if the factor structure is the same across time 
#  [] it looks like there are lots of other types (i.e. weak and strong invariance models) that can compare factor loadings, latent variable means, etc at different times 
#  []its a little weird - there is plenty of info online for getting the syntax, but not for interpretting what the output means - I don't think Mike is going to be teaching us about this so I am on my own 
#  [] CESD is a really solid measure, so I was expecting clean results and it looks like I got them. At this point I am not going to get into interpretting the inferential stuff and the model fit. the diagram is fairly informative for my currently level of understanding. 
```
 
 