---
title: "Making FYP Poster Figures "
---

I am going to make a pedigree, relatedness matrix, interactive relatedness matrix, and fancy survival graph, and then try to get the figure to match my poster color theme. I am also going to try out some cool visualizations of the different kinship links available in the NLSY (though in the poster I end up just using a table for a cleaner look)

```{r message=FALSE, warning=FALSE}
library(NlsyLinks)
library(ggplot2)
library(BGmisc)
library(ggpedigree)
library(tidyr)
library(dplyr)
library(stringr)
library(forcats)
```

## Part 1:  Pedigree of Family 820. 
I picked family 820 because it is one of the larger families available in the data. 
```{r}




```

## Part 2: Survival graph 
```{r}





```

## Part 3: Kinship links visualization 

First, here are the different relatedness values and their frequency
```{r}
r <- c(0, 0.015625,  0.03125,   0.0625,  0.09375, 0.125,   0.1875, 0.25,  0.375, 0.5, 0.75, 1 )
n <- c(1925, 55, 291, 604, 12, 4629, 24,  11294, 647, 22545, 24, 38) 
relatedness <- data.frame(r,n)

relatedness <- relatedness %>%
  mutate(r = as.factor(r))
```

Here is the first crack at it: 
```{r}
ggplot(relatedness, aes(x=reorder(r, -n), y=n)) +
         geom_col(scale_fill_manual = c("gold")) + 
  theme(
    panel.background = element_rect(fill = "#060606"),
    plot.background = element_rect(fill = "#060606"),
    panel.grid = element_blank(),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))
```

This makes it gold, transparent, and saves it directly to my computer so I can upload it into my poster: 
```{r}
plot <- ggplot(relatedness, aes(x = reorder(r, -n), y = n)) +
  geom_col(fill = "#9E7E38") +
  # Add the text labels on top of the bars
  geom_text(
    aes(label = n), 
    vjust = -0.5,     # Adjust vertical position (negative moves it up)
    color = "white",  # Match your dark theme
    size = 7, 
  #  angle = 45,
    # Adjust font size as needed
  family = "Baskerville") +
  #theme_minimal(base_family = "cormorant") +
  theme(text = element_text(family = "Baskerville", color = "white"),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    # Hide the y-axis elements
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    # Style the x-axis for visibility
    axis.text.x = element_text(color = "white", angle = 45,  # Slants the text (45 or 90 are most common)
    hjust = 1, size = 14),
    axis.title.x = element_blank()
  ) +
  # Expand the y-axis slightly so the top label doesn't get cut off
  expand_limits(y = max(relatedness$n) * 1.1) 

ggsave('plot.PNG', plot, bg = 'transparent' )
plot

```

## Part 4: Working on the relatedness matrix and getting the colors right: 
```{r}


```
