---
title: "Making FYP Poster Figures "
---

I am going to make a pedigree, relatedness matrix, interactive relatedness matrix, and fancy survival graph, and then try to get the figure to match my poster color theme. I am also going to try out some cool visualizations of the different kinship links available in the NLSY.

```{r message=FALSE, warning=FALSE, include=FALSE}
library(NlsyLinks)
library(ggplot2)
library(BGmisc)
library(ggpedigree)
library(tidyr)
library(dplyr)
library(stringr)
library(forcats)
library(psych)
library(polycor)
library(tidyverse)
library(devtools)
library(remotes)
library(NlsyLinks)
library(parameters)
library(gtsummary)
library(discord)
library(janitor)
library(lm.beta)
library(matrixStats)
library(gt)
library(OpenMx)
library(ACEsimFit)
#stuff for survival analysis 
library(knitr)
library(survival)
library(ggplot2)
library(tibble)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(condSURV)
library(viridis) 
library(mets)
#new fun package I learned about
library(kableExtra)

# Helper functions
source("~/Documents/Research/Github/risky_gambling/data/miFunctions.R")

#data and cleaning scripts 
source("~/Documents/Research/Github/risky_gambling/data/RiskTaking79/RiskTaking79.R")
#summary(df_RiskTaking79)
source("~/Documents/Research/Github/risky_gambling/data/79Demographics/79Demographics.R")
source("~/Documents/Research/Github/risky_gambling/data/GeneralRisk79/GeneralRisk79.R")
#summary(df_GeneralRisk79)
source("~/Documents/Research/Github/risky_gambling/data/KidRiskBehavior/KidRiskBehavior.R")
#summary(df_KidRiskBehavior)
source("~/Documents/Research/Github/risky_gambling/data/KidsDemographics/KidsDemographics.R")
#summary(df_KidsDemographics)
source('~/Documents/Research/Github/risky_gambling/data/CBRandYBR/CBRandYBR.R')
source('~/Documents/Research/Github/risky_gambling/data/sexID/sexID.R')
source('~/Documents/Research/Github/risky_gambling/data/kidsexID/kidsexID.R')
ID_sex <- rbind(parent_sex, kid_sex)
recode_missing_values <- FALSE
source("~/Documents/Research/Github/risky_gambling/data/Drugs79/Drugs79.R")
source("~/Documents/Research/Github/risky_gambling/data/KidDrugs/KidDrugs.R")
source('~/Documents/Research/Github/risky_gambling/data/age/age.R')
source('~/Documents/Research/Github/risky_gambling/merging_harmonizing_selecting.r')
source('~/Documents/Research/Github/risky_gambling/boring_surv.r')
source('~/Documents/Research/Github/risky_gambling/NLSY_family_creation.r')
source('~/Documents/Research/Github/risky_gambling/BigNLSY_creation.r')
```


## Part 1:  Pedigree of Family 820. 
I picked family 820 because it is one of the larger families available in the data. 

```{r include=FALSE}
nlsy_ped <- BigNLSY |>
  transmute(
    ID = as.integer(ID),
    momID = as.integer(momID),
    dadID = as.integer(dadID),
    sex = sex,
    famID = as.integer(famID),
    lrs = First_Try) #I think lrs is the phenotype so thats what I'm gonna do 

cat("KRSP pedigree:", nrow(nlsy_ped), "individuals,",
  n_distinct(nlsy_ped$famID), "grids\n")

families<- summarizeFamilies(nlsy_ped, famID = "famID")$family_summary |> arrange(desc(count))
```

```{r echo=FALSE}
fam_820 <- nlsy_ped %>%
  filter(famID == 820)
kable(fam_820)
```


To visualize the family, I use the ggpedigree function. I also tried out putting multiple pedigrees next to each other in the same plot. 

```{r echo=FALSE}
mat_820 <- ped2add(fam_820, isChild_method = "partialparent", sparse = FALSE)

ggpedigree::ggpedigree(fam_820,
  personID = "ID",
    config = list(
    code_male = 1, # Here, 1 = male, 0 = female
    sex_color_include = TRUE,
   sex_color_palette = c("black", "#CFB53B")
  ))

fam_521 <- nlsy_ped %>%
  filter(famID == 521) 

fam_711 <- nlsy_ped %>%
  filter(famID == 711) 

df_ped_all <- rbind(fam_820, fam_521, fam_711)


ggpedigree::ggpedigree(df_ped_all,
  personID = "ID",
  famID = "famID",
  config = list(
    code_male = 1, # Here, 1 = male, 0 = female
    sex_color_include = TRUE,
   sex_color_palette = c("black", "#CFB53B")
  ),
  code_male = 1
) +
  facet_wrap(~famID, scales = "free") #+
#  theme_dark() +
#  theme_minimal()
```


For the final poster, I wanted my plot to have a transparent background, and white text, so it pops on my dark background. Then I use ggsave to download it with extra good quality so it doesn't get fuzzy on me. In this case, I am commenting out ggsave so you can see the final thing!

```{r echo = FALSE}
fam_820 <- fam_820 %>%
  rename(personID = "ID")

#plot_820 <- 
  ggPedigree(fam_820,
  famID = "famID",
  personID = "personID",
  config = list(
    focal_fill_personID = 126752035,
    focal_fill_include = TRUE,
     focal_fill_high_color = "gold",
      focal_fill_mid_color = "tan",
      focal_fill_low_color = "#F9F9F7",
    focal_fill_force_zero = TRUE,
    focal_fill_na_value = "white",
    focal_fill_scale_midpoint = 0.25,
    focal_fill_component = "additive",
    focal_fill_method = "gradient",
    focal_fill_n_breaks = NULL,
    focal_fill_legend_title = "Relatedness",
    # "additive",
    sex_color_include = FALSE,
    sex_legend_show = FALSE,
    segment_spouse_color = "white", #I actually think I prefer the white lines 
    segment_sibling_color = "white",
    segment_parent_color = "white",
    segment_offspring_color = "white")) +
  theme(
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))

#ggsave('plot820.PNG', plot_820, bg = "transparent", dpi = 300)

```



## Part 2: Relatedness Matrix of Family 820

First I made an interactive heat map with the default color scheme. 
```{r echo=FALSE, warning=FALSE}
ggRelatednessMatrix(
  mat_820,
  interactive = TRUE,
  config = list(
  #  color_palette = c("white", "skyblue", "darkblue"),
    scale_midpoint = 0.55,
    cluster = TRUE,
    title = "Genetic Relatedness",
   text_size = 6,
    return_widget = TRUE))
```



Next, I manually changed the default settings to get a Wake themed pallete. In the poster, the background is transparent, but that part is commented out here so its more visible here. 

```{r echo=FALSE, warning=FALSE}
source("~/Documents/Research/Github/risky_gambling/buildPLotConfig.R")
source("~/Documents/Research/Github/risky_gambling/ggRelatednessMatrix.R")
source("~/Documents/Research/Github/risky_gambling/defaultPlotConfig_WF.R")

#relate_820 <- 
  ggRelatednessMatrix(mat_820) +
                 # config = default_config)
                 # apply_default_theme == FALSE) +
    theme(text = element_text(family = "Baskerville", color = "white", size = 20),
    panel.background = element_rect(fill = "black", color = NA),
    plot.background = element_rect(fill = "black", color = NA),
    panel.grid = element_blank(),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"),
  #  axis.title = element_blank(),
   # axis.text.x = element_blank(),
   # axis.text.y = element_blank(),
    axis.text.x = element_text(colour = "white", size = 17),
        axis.text.y = element_text(colour = "white", size =17))

#ggsave('relate820.PNG', relate_820, bg = 'transparent', dpi = 300)
```


The Wake theme was not very visually appealing, so I returned to the default colors, and removed the ID axis labels, because they were dominating the visual. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
source("~/Documents/Research/Github/risky_gambling/defaultPlotConfig.R")
relate_820 <- 
  ggRelatednessMatrix(mat_820) +
                 # config = default_config)
                 # apply_default_theme == FALSE) +
    theme(text = element_text(family = "Baskerville", color = "white", size = 20),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"),
    plot.title.position = "plot", #this doesn't work,neither did trying h adjust with plot.title
  #  axis.title = element_blank(),
 #   axis.text.x = element_blank(),
   # axis.text.y = element_blank(),
   axis.text.x = element_text(colour = "white", size = 17),
       axis.text.y = element_text(colour = "white", size =17))

ggsave('relate820.PNG', relate_820, bg = 'transparent', dpi = 300)
```



## Part 3: Survival graph 

I ran a survival analysis and plotted by relatedness group. 

```{r echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#Now I can run a survival analysis comparing groups 
#median age for the 79ers and kids was 16 which is cool 
#first I need to prep some new df format 
df_KidsPairs <- df_linkBigKids %>%
  select(c("First_Try_S1","First_Try_S2","Any_Drug_Use_S1","Any_Drug_Use_S2","FU_Ord_S1","FU_Ord_S2",
           "SubjectTag_S1","SubjectTag_S2","ExtendedID","R","Sex_S1","Sex_S2")) %>%
  mutate(Group = 1)

df_79pairs <- df_linkBigBoy79 %>%
  select(c("First_Try_S1","First_Try_S2","Any_Drug_Use_S1","Any_Drug_Use_S2","FU_Ord_S1","FU_Ord_S2",
           "SubjectTag_S1","SubjectTag_S2","ExtendedID","R","Sex_S1","Sex_S2")) %>%
  mutate(Group =2)

df_linkDrugsSkinny <- rbind(df_KidsPairs, df_79pairs) 

#now I can run the surivval analysis 

old_gold_pallete <- c("#D4AF37", "#D9C589", "#9E7E38", "#7D5F28", 
                   "#5C4018", "#3B2A10", "#B5A68C", "#4A4E41")

survdiff(Surv(First_Try_S1, Any_Drug_Use_S1) ~ R, data = df_linkDrugsSkinny)
survfit(Surv(First_Try_S1, Any_Drug_Use_S1) ~ R, data = df_linkDrugsSkinny)
survfit2(Surv(First_Try_S1, Any_Drug_Use_S1) ~ R, data = df_linkDrugsSkinny) |> 
  ggsurvfit(palette = old_gold_pallete, size =1) +
  labs(
    x = "Years",
    y = "Overall survival probability") +
  add_confidence_interval() +
#  add_risktable() +
   coord_cartesian(xlim = c(0, 30)) +
  theme(text = element_text(family = "Baskerville", color = "white", text.size = 15),
        panel.background = element_rect(fill = "black"),
    plot.background = element_rect(fill = "black"),
    panel.grid = element_blank(),
    legend.text = element_text(color = "white", size =15),
    legend.title = element_text(color = "white", size =15),
    axis.title.x = element_text(colour = "white", size =15),
    axis.text.x = element_text(colour = "white", size =15),
        axis.text.y = element_text(colour = "white", size =15),
    legend.background = element_rect(fill = "black"),
    legend.position = "left") +
  scale_fill_manual(values = old_gold_pallete) +
  scale_color_manual(values = old_gold_pallete)

#ggsave('kapcurves.PNG',plot, dpi = 300)

coxph(Surv(First_Try, Any_Drug_Use) ~ Group, data = df_DrugsSkinny) |> 
 tbl_regression(exp = TRUE) 
```





## Part 4: Kinship links visualization 

First, here are the different relatedness values (r) and their frequency (n).
```{r warning=FALSE, include=FALSE}
r <- c(0, 0.015625,  0.03125,   0.0625,  0.09375, 0.125,   0.1875, 0.25,  0.375, 0.5, 0.75, 1 )
n <- c(1925, 55, 291, 604, 12, 4629, 24,  11294, 647, 22545, 24, 38) 
relatedness <- data.frame(r,n)
```
```{r echo=FALSE, warning=FALSE}
kable(relatedness) %>% kable_styling(full_width = FALSE)
```
```{r warning=FALSE, include=FALSE}
relatedness <- relatedness %>%
  mutate(r = as.factor(r))
```

Here is the first crack at it. Its fine but the colors are blah, the background is black, and the text font doesn't match my poster.   
```{r echo=FALSE, warning=FALSE}
ggplot(relatedness, aes(x=reorder(r, -n), y=n)) +
         geom_col(scale_fill_manual = c("gold")) + 
  theme(
    panel.background = element_rect(fill = "#060606"),
    plot.background = element_rect(fill = "#060606"),
    panel.grid = element_blank(),
    legend.text = element_text(color = "white"),
    legend.title = element_text(color = "white"))
```

This makes it gold, transparent, and saves it directly to my computer so I can upload it into my poster: 
```{r echo=FALSE, message=FALSE, warning=FALSE}
plot <- ggplot(relatedness, aes(x = reorder(r, -n), y = n)) +
  geom_col(fill = "#9E7E38") +
  # Add the text labels on top of the bars
  geom_text(
    aes(label = n), 
    vjust = -0.5,     # Adjust vertical position (negative moves it up)
    color = "white",  # Match your dark theme
    size = 7, 
  #  angle = 45,
    # Adjust font size as needed
  family = "Baskerville") +
  #theme_minimal(base_family = "cormorant") +
  theme(text = element_text(family = "Baskerville", color = "white"),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA),
    panel.grid = element_blank(),
    # Hide the y-axis elements
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    # Style the x-axis for visibility
    axis.text.x = element_text(color = "white", angle = 45,  # Slants the text (45 or 90 are most common)
    hjust = 1, size = 14),
    axis.title.x = element_blank()
  ) +
  # Expand the y-axis slightly so the top label doesn't get cut off
  expand_limits(y = max(relatedness$n) * 1.1) 

ggsave('plot.PNG', plot, bg = 'transparent' )
plot

```

### Conclusion
In this portfolio, I've applied various graphical techniques to illustrate some of the sample characteristics and results from my study. Thank you for tuning in! 
