---
title: "Airline Loyalty Data - Data Wrangling Practice"
---

> To wrap up this portfolio, I found some data online (https://mavenanalytics.io/data-playground/airline-loyalty-program). Hopefully this is going to be a fun feel good activity. This data is clean and ready to go, and the place where I downloaded it from even gave some ideas for what to explore!


## Lets check out the data first: 
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(haven)
library(knitr)
library(kableExtra)
library(tidyr)
library(dplyr)
library(ggplot2)
library(ggalt)

activity <- read.csv("Airline+Loyalty+Program/Customer Flight Activity.csv")
history <- read.csv("Airline+Loyalty+Program/Customer Loyalty History.csv")
calendar <- read.csv("Airline+Loyalty+Program/Calendar.csv")
dictionary <- read.csv("Airline+Loyalty+Program/Airline Loyalty Data Dictionary.csv")
kable(head(dictionary), caption = "Codebook")
kable(head(activity), caption = "Travel Stats of Loyalty Members")
history_table <-head(history)
kable(history_table[,1:8], caption = "Loyalty Member Demographics") 
kable(history_table[,9:16]) 
```

There was a promotional running from Feb - Apr 2018.
We have data on:
* How long someone has been part of the flight loyalty program 
* How much they have flown with the airline 
* Points accumulated and redeemed 
* Marital status, education, income, where the person lives, etc 

### Analysis Questions: 

* What impact did the campaign have on loyalty program memberships (gross / net)?
* Was the campaign adoption more successful for certain demographics of loyalty members? 
* What impact did the campaign have on booked flights during the summer?


### The campaign relates to a spike in enrollment

```{r echo=FALSE, message=FALSE, warning=FALSE}
history <- history %>%
  filter(Enrollment.Year > 2015)

ggplot(history, mapping = aes(x=Enrollment.Month, fill=factor(Enrollment.Month))) + 
  geom_histogram() + 
  facet_grid(~Enrollment.Year) +
   scale_fill_manual(values = c("grey","red","red","red","grey","grey","grey","grey","grey","grey","grey","grey"))+
  theme_minimal() + 
  labs(title = "Monthly Enrollments in the Loyalty Program from 2016 to 2018",
       subtitle = "Jumps in Enrollment Correspond with the Feb - April 2018 Campaign",
       x = "Month",
       y = "Number of Enrollments",
       fill = "Month") + 
  theme(axis.text.x = element_blank())

```

Dumbbell chart looking at 2017 monthly enrollments versus 2018. This one really magnifies the differences to make it seem like there were huge changes between 2017 and 2018, even after the campaign had ended . 

```{r echo=FALSE, message=FALSE, warning=FALSE}
dumbell <- history %>% filter(Enrollment.Year > 2016) %>% select(c("Enrollment.Year","Enrollment.Month", "Loyalty.Number"))

dumbell <- data.frame(table(dumbell$Enrollment.Month, dumbell$Enrollment.Year))

dumbell <- dumbell %>% rename(Month = "Var1", Year = "Var2")

dumbell <- dumbell %>% pivot_wider(names_from = "Year",
                        values_from = "Freq",
                        id_cols = "Month",
                        names_prefix = "year_")
```
```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(dumbell, mapping = aes(x=year_2017, xend = year_2018, y = Month, group = Month)) +
  geom_dumbbell(colour = "lightblue",
                colour_xend = "#2A52BE",
                size = 2,
                dot_guide = TRUE,
                dot_guide_size = .15,
                dot_guide_colour = "grey") +
  coord_flip() +
  labs(title = "Changes in Monthly Enrollment from 2017 to 2018",
       x = "Month",
       y = "Number of New Enrollments") +
  theme_classic() 

```

### There do not appear to be gender differences in enrollments
```{r echo=FALSE, message=FALSE, warning=FALSE}

history <- history %>%
  filter(Enrollment.Year > 2017)

ggplot(history, mapping = aes(x=Enrollment.Month, fill=factor(Enrollment.Month))) + 
  geom_histogram() + 
  facet_grid(~Gender) +
   scale_fill_manual(values = c("grey","red","red","red","grey","grey","grey","grey","grey","grey","grey","grey"))+
  theme_minimal()+ 
  labs(title = "Monthly Enrollments in the Loyalty Program in 2018",
       subtitle = "Jumps in Enrollment Correspond with the Feb - April 2018 Campaign",
       x = "Month",
       y = "Number of Enrollments",
       fill = "Month") + 
  theme(axis.text.x = element_blank())

```

### We could use this chart to say that enrollments are lower for more educated people (but I imagine that if you make enrollment count proportional to the number of people with that degree, we would see something different)
```{r echo=FALSE, message=FALSE, warning=FALSE}

history <- history %>%
  mutate(group = factor(Education, levels = c("High School or Below","College","Bachelor","Master","Doctor")))

ggplot(history, mapping = aes(x=Enrollment.Month, fill=factor(Enrollment.Month))) + 
  geom_histogram() + 
  facet_grid(~group) +
   scale_fill_manual(values = c("grey","red","red","red","grey","grey","grey","grey","grey","grey","grey","grey"))+
  theme_minimal() + 
  labs(title = "Monthly Enrollments in the Loyalty Program in 2018",
       subtitle = "Jumps in Enrollment Correspond with the Feb - April 2018 Campaign",
       x = "Month",
       y = "Number of Enrollments",
       fill = "Month") + 
  theme(axis.text.x = element_blank())

```

## There was a big impact on booked flights over the summer

```{r echo=FALSE, message=FALSE, warning=FALSE}
act <- activity %>% select(c("Loyalty.Number","Year","Month","Total.Flights", "Distance")) %>%
  pivot_wider(names_from = "Year",
                        values_from = c("Total.Flights","Distance"),
                        values_fn = sum,
                        id_cols = "Month",
                        names_prefix = "year_") 

act <- act %>% pivot_longer(cols = c("Total.Flights_year_2018","Total.Flights_year_2017","Distance_year_2018","Distance_year_2017"),
                     names_to = c("Act_Type","Year"),
                     names_sep = "_year_"
                    )

act %>% filter(Act_Type == "Total.Flights") %>% 
ggplot(act, mapping = aes(x=Month, y=value, color = Year)) +
  geom_line() +
  theme_classic() +
  labs(title = "Total Flights in 2017 and 2018",
       x = "Month",
       y= "Total Flights",
       color = "Year")


act %>% filter(Act_Type == "Distance") %>% 
  mutate(dist_mil = value / 1000000) %>%
ggplot(act, mapping = aes(x=Month, y=dist_mil, color = Year)) +
  geom_line() +
  theme_classic() +
  labs(title = "Total Flight Miles in 2017 and 2018",
       x = "Month",
       y= "Total Miles (in Millions)",
       color = "Year") +
  theme(axis.text.y = element_text(angle = 45),
        axis.text.x = element_text(angle =45),
        text = element_text(family = "Baskerville", size = 15))

```

### You'll see that even though the campaign ended iin April 2018, we don't see the effect of the campaign on bookings until May-ish. Between the gross enrollments and the activity plots, it appears that the campaign led people to enroll a lot more than usual between feb and april, and this led to a spike in bookings for the summer. 


# The last thing I am going to do is try to map the members using either their zip codes or cities, following the La Quinta Lab 

I am using the mapcan package to get the outline / map of Canada 

Here is the US, just because. I really like the usmap package - cool stuff!
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(usmap)
library(ggthemes) # for the theme_map

plot_usmap(
  regions = c("states", "state", "counties", "county"),
  include = c(),
  exclude = c(),
  data = data.frame(),
  values = "values",
  theme = theme_map(),
  labels = FALSE,
  label_color = "black",
  data_year = NULL
)

```

Canada initial map 
```{r echo=FALSE, message=FALSE, warning=FALSE}
#haha never mind we only have canada

library(mapcan)
canada_map <- mapcan(boundaries = province, type = standard)

ggplot(canada_map, aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black") +
  theme_mapcan() +
  coord_fixed() +
  labs(title = "Loyalty Members in Canada")


```

It turns out that there is not an easy R package (like zipcodeR) for canadian postal codes long and lat, so I am aborting this mission. 
